{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-11 thesis/overview.tex
rearrange some overview mc prose

@@

Model checking \cite{verisoft} is a testing technique for systematically exploring the possible thread interleavings of a concurrent program.
A model checker executes the program repeatedly, each time according to a new thread interleaving, until the state space (or the CPU budget) is exhausted.
During each execution, it forces threads to execute serially, thereby confining the program's nondeterminism to controlled thread switches.
Using a single iteration of the {\tt x++;} loop from Figure~\ref{fig:concurrency-bug} as an example,
Figure~\ref{fig:tree}(a) shows all possible execution interleavings of the compiled code between 2 threads.

@@

Some model checkers explicitly store the set of visited program states as a means of identifying equivalent interleavings \cite{spin}.
This approach is called {\em stateful} model checking.
In this thesis, I focus on {\em stateless} model checking,
which instead analyzes the sequence of execution events to avoid a prohibitive memory footprint.
Henceforth I will abbreviate ``stateless model checking'' simply as ``model checking'' for brevity.

@@

During execution, a model checker identifies a subset of the program's operations as ``interesting'', i.e.,
where interrupting the current thread to run a different one is likely to produce different behaviour.
These so-called {\em preemption points} may be identified by any combination of human intuition and machine analysis.
Typical preemption points include the boundaries of synchronization APIs (e.g., {\tt mutex\_lock}) or accesses to shared variables.
Considering that at each preemption point multiple threads exist as options to run next,
the set of possible ways to execute the program can be viewed as a tree.
Figure~\ref{fig:tree}(b) shows a visualization of the corresponding tree from our example program.

@@

@@

% concurrency bug
% race condition: "a confusing term which in some circles means concurrency bug and others data race, which we will avoid"
% data race
% atomicity violation (what even is this.)
% thread
% transition
% conflict
% independent: "see conflict"
% state space
% interleaving
% DPOR
% estimation (WBE, RE)
% landslide
% quicksand
% iterative deepening
% MC, stateless MC
% TA: teaching assistant
% P2 (note abotu psu - dosent call it p2, but we call it that for both schools here
% happens before, all 3 versions - DPOR, LHB, and PHB
% userspace, kernelspace

>>>>

Model checking \cite{verisoft} is a testing technique for systematically exploring the possible thread interleavings of a concurrent program.
A model checker executes the program repeatedly, each time according to a new thread interleaving, until all interleavings have been tested or the CPU budget is exhausted.
During each execution, it forces threads to execute serially, thereby confining the program's nondeterminism to scheduler thread switches.
It then controls the scheduling decisions
%, choosing a different thread from one chosen before at a certain execution states
to guarantee a unique interleaving is tested each iteration.

@@

To understand what it means to exhaustively test all possible thread interleavings,
one must define the possible execution sequences as a finite {\em state space}.
To visualize this,
using a single iteration of the {\tt x++;} loop from Figure~\ref{fig:concurrency-bug} as an example,
with {\tt x++;} expanded into its three corresponding pseudo-assembly instructions,
Figure~\ref{fig:tree}(a) shows all possible execution interleavings
%of the compiled code
between 2 threads.

@@

@@

During execution, a model checker identifies a subset of the program's operations as ``interesting'', i.e.,
where interrupting the current thread to run a different one is likely to produce different behaviour.
These so-called {\em preemption points} may be identified by any combination of human intuition and machine analysis.
Typical preemption points include the boundaries of synchronization APIs (e.g., {\tt mutex\_lock}) or accesses to shared variables.
Considering that at each preemption point multiple threads exist as options to run next,
the set of possible ways to execute the program can be viewed as a tree.
Figure~\ref{fig:tree}(b) shows a visualization of the corresponding tree from our example program,
using each pseudo-assembly instruction as a preemption point.

@@

Some model checkers explicitly store the set of visited program states as a means of identifying equivalent interleavings \cite{spin}.
From the perspective of such tools, state spaces such as these wherein equivalent states may be reached by multiple paths
are represented as a directed acyclic graph (DAG) instead of as a tree.
This approach is called {\em stateful} model checking.
In this thesis, I focus on {\em stateless} model checking (and execution trees, not DAGs),
which instead analyzes the sequence of execution events to avoid a prohibitive memory footprint.
Henceforth I will abbreviate ``stateless model checking'' simply as ``model checking'' for brevity.
Also, the term ``state space'' was originally coined to refer to the stateful approach's emphasis on the DAG's nodes
(i.e., program states);
while stateless checkers emphasize the tree's branches (i.e., execution sequences) instead,
I will continue to use ``state space'' for consistency with prior work.

@@

% concurrency bug
% race condition: "a confusing term which in some circles means concurrency bug and others data race, which we will avoid"
% data race
% atomicity violation (what even is this.)
% thread
% transition
% conflict
% independent: "see conflict"
% state space
% interleaving
% DPOR
% estimation (WBE, RE)
% ETA
% landslide
% quicksand
% iterative deepening
% MC, stateless MC
% TA: teaching assistant
% P2 (note abotu psu - dosent call it p2, but we call it that for both schools here
% happens before, all 3 versions - DPOR, LHB, and PHB
% userspace, kernelspace
